---
execute:
  echo: true
  message: false
  warning: false
format:
  html:
    fig-width: 6
    fig-height: 4
    out-width: 60%
    embed-resources: true
---
# Results


```{r}
# Import libraries
# install.packages("choroplethr")
# install.packages("RColorBrewer")
# install.packages("GGally")
# install.packages("devtools")
# install.packages("remotes")
# install.packages("choroplethrMaps")
# devtools::install_github("timelyportfolio/parcoords")
# install.packages("ggpmisc")

library(ggpmisc)
library(tidyverse)
library(data.table)
library(vcd)
library(choroplethrMaps)
library(GGally)
library(parcoords)
library(RColorBrewer)
library(choroplethr)
library(lubridate)
library(stringr)

```



```{r}
#Import starting dataset and filter for vaccine hesitance indicator and monthly samples


# Load vaccine hesitancy data
covid <- fread("../covid_dataset/covid.csv")

# Filter for Monthly sample period and vaccine hesitance responses
covid_hes <- covid |>
  filter(`Time Type` == "Monthly",
         `Indicator Category` == "Probably or definitely will not get vaccinated")

```



```{r}
#Prep data for analysis. The following chunk truncates the names of population segments for data visualization purposes

# Create vector of shortened group names
group_name_shortening <- c("Age" = "Age",
                           "Age by race/ethnicity" = "Age.Race",
                           "All adults 18+" = "18+ Adults",
                           "Born in the U.S." = "US.Born",
                           "Concern about getting COVID-19 disease" = "Covid.Concern",
                           "Confidence in COVID-19 vaccine safety" = "Vaccine.Confiden",
                           "Confidence that COVID-19 vaccine is important" = "Vaccine.Imprtnt",
                           "COVID-19 vaccination status of friends and family" = "Fr.Fam.Vaccd",
                           "COVID-19 vaccine information seeking frequency in past month" = "Vacc.Info.Seek",
                           "Disability status (any)" = "Disability",
                           "Essential worker group" = "Ess.Wrkr.grp",
                           "Ever had COVID-19 disease (self-report)" = "Past.Covid",
                           "Gender identity" = "G.Identity",
                           "Have a regular physician or provider for primary care" = "Prim.Care.Providr.Status",
                           "Health condition associated with higher risk for COVID-19 (any)" = "Comorbidity",
                           "Healthcare provider recommended I get a COVID-19 booster" = "Dr.Recommend.Boostr",
                           "Healthcare provider recommended I get a COVID-19 vaccine" = "Dr.Recommend.Vacc",
                           "Insurance status" = "Insurance",
                           "Language of interview" = "Language",
                           "Mask-wearing frequency" = "Mask.Use",
                           "Mental health status" = "Mental.Health",
                           "Metropolitan statistical area" = "Rural.Urban",
                            "Political leaning of county of residence" = "Political",
                           "Poverty status" = "Pov.Status",
                           "Pregnancy status (females age 18 – 49 years)" = "Pregnancy",
                           "Race/Ethnicity (4 level)" = "Race.4lvl",
                           "Race/Ethnicity (7 level)" = "Race.7lvl",
                           "Received a flu vaccination" = "Flu.Vacc",
                           "Received non-COVID-19 vaccine(s) in past two years" = "Past.Vaccin",
                           "Self-perceived equity in seeking healthcare compared to other races/ethnicities" = "Racial.Equity",
                           "Sex" = "Sex",
                           "Sexual orientation" = "S.Orientation",
                           "Social Vulnerability Index (SVI) of county of residence" = "SVI",
                           "Speaks a language other than English at home" = "Non.Englsh.Language",
                           "Work or school requires COVID-19 vaccination" = "Work.School")

# Construct dataframe from vector
group_name_shortened <- data.frame(`Group Name` = names(group_name_shortening), shortened_group = group_name_shortening, stringsAsFactors = FALSE)|>
  dplyr::rename(`Group Name` = Group.Name)

# Create vector of shortened group category names
group_cat_shortening <- c("18 – 49 years" = "18 – 49",
                          "50 – 64 years" = "50 – 64",
                          "65+ years" = "65+",
                          "18 – 49 years, Black, non-Hispanic" = "18 – 49, Black",
                          "18 – 49 years, Hispanic" = "18 – 49, Hispanic",
                          "18 – 49 years, Other or multiple races, non-Hispanic" = "18 – 49, Other",
                          "18 – 49 years, White, non-Hispanic" = "18 – 49, White",
                          "50 – 64 years, Black, non-Hispanic" = "50 – 64, Black",
                          "50 – 64 years, Hispanic" = "50 – 64, Hispanic",
                          "50 – 64 years, Other or multiple races, non-Hispanic" = "50 – 64, Other",
                          "50 – 64 years, White, non-Hispanic" = "50 – 64, White",
                          "65+ years, Black, non-Hispanic" = "65+, Black",
                          "65+ years, Hispanic" = "65+, Hispanic",
                          "65+ years, Other or multiple races, non-Hispanic" = "65+, Other",
                          "65+ years, White, non-Hispanic" = "65+, White",
                          "Born in the U.S." = "US Born",
                          "Not born in the U.S." = "Not US Born",
                          "Many or almost all family or friends have received vaccine" = "Many",
                          "Some or no family or friends have been vaccinated" = "No_some","A little or not at all concerned" = "A little_not",
                          "Very or moderately concerned" = "Very_Moderately",
                          "Somewhat or not at all safe for me" = "Not at all_Somewhat",
                          "Very or completely safe for me" = "Very",
                          "A little or not at all important to protect me from COVID-19" = "A little_not at all",
                          "Somewhat or very important to protect me from COVID-19" = "Somewhat_very",
                          "No" = "No",
                          "Yes" = "Yes",
                          "Cisgender" = "Cisgender",
                          "Don't know/refused" = "Dont know",
                          "Transgender/Non-binary" = "TG/Non bin",
                          "Insured" = "Insured",
                          "Not insured" = "Not insured",
                          "English" = "English",
                          "Other" = "Other",
                          "Spanish" = "Spanish",
                          "Rural" = "Rural",
                          "Suburban" = "Suburban",
                          "Urban" = "Urban",
                          "Democrat-leaning" = "Democrat",
                          "Not Democrat-leaning or Republican-leaning" = "None",
                          "Republican-leaning" = "Republican",
                          "Above poverty, income <$75k" = "<$75k",
                          "Above poverty, income >=$75k" = ">=$75k",
                          "Below poverty" = "Below poverty",
                          "Unknown income" = "Unknown",
                          "Breastfeeding" = "Breastfeeding",
                          "Not pregnant, trying to get pregnant, or breastfeeding" = "Not pregnant",
                          "Pregnant" = "Pregnant",
                          "Trying to get pregnant" = "Trying to get pregnant",
                          "American Indian/Alaska Native, non-Hispanic" = "Native",
                          "Asian, non-Hispanic" = "Asian",
                          "Black, non-Hispanic" = "Black",
                          "Hispanic" = "Hispanic",
                          "Native Hawaiian/Pacific Islander, non-Hispanic" = "Pacific",
                          "Other or multiple races, non-Hispanic" = "Other",
                          "White, non-Hispanic" = "White",
                          "Female" = "Female",
                          "Male" = "Male",
                          "Gay/Lesbian/Bisexual/Other" = "LGB",
                          "Heterosexual/Straight" = "Hetero",
                          "High SVI" = "High",
                          "Low SVI" = "Low",
                          "Moderate SVI" = "Moderate",
                          "SVI unknown" = "Unknown",
                          "Not applicable/unemployed" = "Unemployed",
                          "All adults age 18+ years" = "18+ Adults",
                          "Essential healthcare" = "Ess.Healthcare",
                          "Other essential worker" = "Ess.Other",
                          "Other frontline worker" = "Frontline.Other",
                          "Persons who are not essential workers" = "Not.Ess.Wrkr",
                          "School and childcare" = "School.Childcare",
                          "Better than other races or ethnicities" = "Better",
                          "Same as other races or ethnicities" = "Equal",
                          "Worse than other races or ethnicities" = "Worse",
                          "Did not receive flu vaccination" = "No",
                          "Received a flu vaccination" = "Yes",
                          "Political leaning unknown" = "Unknown",
                          "Often or sometimes" = "Often.Sometimes",
                          "Rarely or never" = "Rarely.Never",
                          "Always or often wears mask" = "Always.Often",
                          "Sometimes, rarely, or never wears mask" = "Occasional.Never",
                          "Excellent, very good, or good" = "Excellent.Good",
                          "Fair or poor" = "Fair.Poor")

# Construct dataframe from vector
group_cat_shortened <- data.frame(`Group Category` = names(group_cat_shortening), shortened_cat = group_cat_shortening, stringsAsFactors = FALSE) |>
  dplyr::rename(`Group Category` = Group.Category)

# Join shortened group names and categories to master table
covid_hes_prepped <- left_join(left_join(covid_hes, group_name_shortened,
                                         by = "Group Name"),
                                         group_cat_shortened, by = "Group Category")


# Create combinations of group names and categories via paste function
# Include string wrapped version of text for visualization purposes
covid_hes_prepped <- covid_hes_prepped |>
  mutate(concat = paste(shortened_group, shortened_cat, sep = " - "),
         concat_wrap = str_wrap(concat, width = 10))


```


## Part 1: Study which segments of the population exhibited greatest covid vaccine hesitancy following the deployment of mass vaccinations in the US in early 2021. 

The national average COVID hesitance rate reported during the first quarter of the data collection period (April - July 2021) was 17.7%. In Part 1, we will break down this percentage across U.S. geographies, identifying the states with the highest and lowest hesitancy rates. We will then analyze the types of population demographics (e.g., age, vaccination history, ethnicity, etc.) that had a strong association with COVID vaccine hesitancy.


To obtain an initial baseline of vaccine hesitance rates, we developed the chunk of code below, which revealed that the nationwide hesitance rate from April 2021 to July 2021 was 17.7% nationwide. In other words, 17.7% of those surveyed after the mass deployment of vaccinations responded that they would probably or definitely not get vaccinated.
```{r}

# Store values of months that will be combined to analyze hesitancy when vaccines were initially released in 2021
periods_to_combine_2021 <- c("April 22 - May 29","May 30 - June 26","June 27 - July 31")

# Calculate overall nationwide hesitancy
national_hesitance_rate <- covid_hes_prepped |>
  filter(`Year` == 2021,
         `Time Period` %in% periods_to_combine_2021,
         `Group Category`== "All adults age 18+ years",
         Geography == "National") |>
  mutate(n_hesitant = round(`Estimate (%)`/100 * `Sample Size`,0)) |>
  group_by(Geography) |>
  summarise(`Sample Size` = sum(`Sample Size`), n_hesitant = sum(n_hesitant)) |>
  mutate(`Estimate (%)` = n_hesitant / `Sample Size` * 100,
         `Time Period` = "2021 analysis period",
         Year = 2021) 

```

We began our analysis of variation in hesitancy rates by calculating hesitancy during the April-July 21' timeframe by state. As illustrated in the map below, states such as Hawaii, New York, and Vermont exhibited the lowest hesitance, while other states like Wyoming, North Dakota, and West Virginia exhibited the highest hesitance.

```{r}

# Calculate early 2021 hesitance by state
covid_2021_analysis_allstates <- covid_hes_prepped |>
  filter(`Year` == 2021,
         `Time Period` %in% periods_to_combine_2021,
         `Group Category`== "All adults age 18+ years",
         `Geography Type` %in% c("Jurisdictional Estimates"),
         !grepl("-", Geography)) |>
  mutate(n_hesitant = round(`Estimate (%)`/100 * `Sample Size`,0)) |>
  group_by(Geography) |>
  summarise(`Sample Size` = sum(`Sample Size`),
            n_hesitant = sum(n_hesitant)) |>
  mutate(`Estimate (%)` = n_hesitant / `Sample Size` * 100,
         `Time Period` = "2021 analysis period",
         Year = 2021)

# Prep dataframe to be used in the state_choropleth function
covid_2021_analysis_allstates_map <- covid_2021_analysis_allstates |>
  select(Geography, `Estimate (%)`) |>
  transmute(region = tolower(Geography), value = `Estimate (%)`)


# Display state heatmap of hesitance rates
state_choropleth(covid_2021_analysis_allstates_map,
                 title = "   April-July 2021 COVID-19 Vaccine Hesitance Rates by State",
                 legend = "Hesitancy Rates (%) ")


```

The histogram and boxplot below show that there are no clear outlier states. The code we developed also searches for outlier values using the lower and upper bounds of the distribution (lower bound = Q1 - 1.5xIQR, upper bound = Q3 + 1.5xIQR). While no outliers were identified, we will see how hesitance rates within states can vary significantly from one population segment to another.  

```{r}

# Plot histogram of state hesitance rates
ggplot(covid_2021_analysis_allstates, aes(x = `Estimate (%)`)) +
  geom_histogram(bins = 8, fill = "blue", color = "black", alpha = 0.7) +
  geom_boxplot() +
  labs(title = "State distribution of hesitance", x = "Hesitance %", y = "No. of states")

# Search for outliers
n_outliers <- covid_2021_analysis_allstates |>
  mutate(lower_bound = quantile(`Estimate (%)`, 0.25) - 1.5*IQR(`Estimate (%)`),
         upper_bound = quantile(`Estimate (%)`, 0.75) + 1.5*IQR(`Estimate (%)`)) |>
  filter(`Estimate (%)` < lower_bound & `Estimate (%)` > upper_bound)

```

Before diving into each state further, we examined the population segments with relatively higher vaccine hesitance rates from April-July 2021. Nationally, it can be seen that the 10 segments displayed below demonstrated the highest hesitance. Each segment represents a group of individuals who provided a particular response. For instance, "Fr.Fam.Vaccd - No_some" represents individuals who responded "No or some" to whether their friends and/or family are vaccinated.

```{r}


# Create vector of all group names as represented in the data
group_name_toloop <- c("Sex","Age","Race/Ethnicity (7 level)","Age by race/ethnicity","Sexual orientation","Gender identity","Metropolitan statistical area","Born in the U.S.", "Language of interview","Poverty status", "Insurance status","Social Vulnerability Index (SVI) of county of residence", "Political leaning of county of residence", "Received non-COVID-19 vaccine(s) in past two years", "Health condition associated with higher risk for COVID-19 (any)","Disability status (any)", "Pregnancy status (females age 18 – 49 years)", "Ever had COVID-19 disease (self-report)","Concern about getting COVID-19 disease", "Confidence in COVID-19 vaccine safety",  "Confidence that COVID-19 vaccine is important", "Healthcare provider recommended I get a COVID-19 vaccine",  "COVID-19 vaccination status of friends and family","Work or school requires COVID-19 vaccination" )


group_name_toloop_short <- lapply(group_name_toloop, function(x) ifelse(x %in% names(group_name_shortening), group_name_shortening[[x]], x))
  

# Calculate vaccine hesitance by population segment
covid_2021_analysis <- covid |>
  filter(`Year` == 2021, `Time Period` %in% periods_to_combine_2021,
         `Group Name` %in% group_name_toloop,
         `Geography Type` %in% c("Jurisdictional Estimates",
                                 "National Estimates"),
         `Indicator Category` == "Probably or definitely will not get vaccinated",
         !grepl("-", Geography)) |>
  mutate(n_hesitant = round(`Estimate (%)`/100 * `Sample Size`,0)) |>
  group_by(pick(c(1:6))) |>
  summarise(`Sample Size` = sum(`Sample Size`),
            n_hesitant = sum(n_hesitant)) |>
  filter(Geography == "National") |>
  mutate(`Estimate (%)` = n_hesitant / `Sample Size` * 100,
         `Time Period` = "2021 analysis period",Year = 2021)


# Join shortened group names and categories to master table
covid_2021_analysis <- left_join(left_join(covid_2021_analysis, group_name_shortened,
                                         by = "Group Name"),
                                         group_cat_shortened, by = "Group Category")


# Create combinations of group names and categories via paste function
# Include string wrapped version of text for visualization purposes
covid_2021_analysis <- covid_2021_analysis |>
  mutate(concat = paste(shortened_group, shortened_cat, sep = " - "),
         concat_wrap = str_wrap(concat, width = 10),
         `Group Name` = shortened_group,
         `Group Category` = shortened_cat) |>
  select(-c(shortened_group, shortened_cat))


# Visualize the 10 categories with the highest hesitance rates
covid_2021_analysis |> ungroup() |> top_n(10, `Estimate (%)`) |>
  ggplot(aes(x = fct_reorder(concat_wrap, -(`Estimate (%)`)),
             y = `Estimate (%)`)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 categories with the highest hesitance in 2021",
       x = "Group categories", y = "2021 hesitance (%)") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1,size=6),
        axis.text.y = element_text(size = 8),
        axis.title = element_text(size = 8),
        plot.title = element_text(size = 12))

# Store the categories with the highest hesitance rates in a single-column dataframe
top_10_highest_hesitance_cat_list <- covid_2021_analysis |>
  ungroup() |>
  top_n(10, `Estimate (%)`) |> select(concat)

```


To dive deeper into the variations in vaccine hesitance by state, we focused on the 10 most and least hesitant states and explored how their respective hesitance rates differed by population segment. Rather than considering all population segments, we narrowed the groups to those that, nationally, had the highest hesitance (see graph above).

To facilitate our investigation, we created a set of parallel coordinate plots for the group of 10 most hesitant states and the group of 10 least hesitant states. In both cases, the hesitance rates were not uniform across all population segments. In particular, the group of individuals who responded "A little or not at all" to whether vaccines are important to them showed remarkably higher hesitance rates at the state level.

[to review]
Another question we looked at via interactive parallel coordinates plots is whether states with higher overall hesitance rates would consistently have higher rates at the population segment level. Conversely, we looked at whether states with lower overall hesitance rates had consistently lower rates relative to other states at the population segment level. The plots created revealed that no clear pattern existed for knowing whether a top 10 hesitant state would necessarily be in the top 10 for all population segments. For example, California had the highest hesitance among white individuals aged 18-49 but was not in the top 9 states for hesitance among those who have not previously been vaccinated (see segment = "Past.Vaccin - No").

```{r}

# Calculate hesitance over three months in 2021 by state and population segment
# Clean state column
state_group_analysis <- covid |>
  filter(`Year` == 2021,
         `Time Period` %in% periods_to_combine_2021,
         `Geography Type` == "Jurisdictional Estimates",
         `Indicator Category` == "Probably or definitely will not get vaccinated",
         !grepl("-", Geography)) |>
  mutate(n_hesitant = round(`Estimate (%)`/100 * `Sample Size`, 0)) |>
  group_by(Geography, `Group Name`, `Group Category`) |>
  summarise(`Sample Size` = sum(`Sample Size`),
            n_hesitant = sum(n_hesitant)) |>
  mutate(`Estimate (%)` = n_hesitant / `Sample Size` * 100,
         `Time Period` = "2021 analysis period",
         Year = 2021) |>
  ungroup()


# Shorten group names
state_group_analysis <- state_group_analysis |>
  mutate_at(vars(`Group Name`), ~recode(., !!!group_name_shortening))

state_group_analysis <- state_group_analysis |>
  mutate_at(vars(`Group Category`), ~recode(., !!!group_cat_shortening))

state_group_analysis <- state_group_analysis |> mutate(concat = paste(`Group Name`, `Group Category`, sep = " - "))

state_group_analysis$concat_wrap <- str_wrap(state_group_analysis$concat, width = 10)


# Filter groups for top most hesitant
# Pivot wider to convert row-level categories to columns containing hesitancy rates
state_group_analysis_pivot <- state_group_analysis |>
  filter(concat %in% top_10_highest_hesitance_cat_list$concat) |>
  select(Geography, `Estimate (%)`, concat_wrap) |>
  pivot_wider(id_cols = Geography, names_from = "concat_wrap",
              values_from = "Estimate (%)")

# Pregnancy - pregnant is not available at the state level during the filtered timeframe
state_not_in_top_10 <- top_10_highest_hesitance_cat_list |>
  filter(!concat %in% state_group_analysis$concat)

pregnancy_check <- covid |>
  filter(`Group Category` == "Pregnant") |>
  distinct(`Geography Type`, Geography, `Group Name`, `Group Category`,
           `Time Period`, Year)

# Identify top most hesitant states
top_10_hesistant_states <- covid_2021_analysis_allstates |>
  arrange(desc(`Estimate (%)`)) |>
  filter(!Geography %in% c("Guam","Puerto Rico")) |>
  slice(1:10)
  
# Identify top least hesitant states
bottom_10_hesistant_states <- covid_2021_analysis_allstates |>
  arrange(`Estimate (%)`)|>
  filter(!Geography %in% c("Guam","Puerto Rico")) |>
  slice(1:10)

# Filter pivoted data on most hesitant states
top_10_parallel_analaysis <- state_group_analysis_pivot |>
  filter(Geography %in% top_10_hesistant_states$Geography)


# Display parallel coordinates plot for most hesitant states
ggparcoord(top_10_parallel_analaysis |> select(where(~!any(is.na(.)))), columns = c(2:8),
           scale = "globalminmax",
           alpha = 0.4) +
    labs(title = "Top categories with the highest hesitance in 2021 (10 highest hesitance states)",
       x = "Group categories", y = "2021 hesitance (%)")


# Display interactive version
top_10_parallel_analaysis |> select(where(~!any(is.na(.))))|>
  parcoords(rownames = F, brushMode = "1D-axes", reorderable = T, queue = T)


# Filter pivoted data on least hesitant states
bottom_10_parallel_analaysis <- state_group_analysis_pivot |>
  filter(Geography %in% bottom_10_hesistant_states$Geography)


# Display parallel coordinates plot for least hesitant states
ggparcoord(bottom_10_parallel_analaysis |> select(where(~!any(is.na(.)))), 
           columns = c(2:7),   
           scale = "globalminmax",
           alpha = 0.4) +
    labs(title = "Top categories with the highest hesitance in 2021 (10 lowest hesitance states)",
       x = "Group categories", y = "2021 hesitance (%)")


# Display interactive version
bottom_10_parallel_analaysis |>
  select(where(~!any(is.na(.))))|>
  parcoords(rownames = F, brushMode = "1D-axes", reorderable = T, queue = T)

```


Given the inconsistent trends seen across states and population categories with the highest hesitance, we proceeded to conduct chi squared tests on every group name defined within the scope of this exercise to determine those with the highest statistical significance at a national level.

```{r}

get_chi_sq <- function(df, geography, time_period, year, group_name, hesitant_group) {
  
  temp <- df |> filter(Geography == geography,
                       `Indicator Name`== "Vaccination uptake and intention",
                       `Time Period` == time_period,
                       Year == year,
                       `Group Name`== group_name) |>
    mutate(hesitant = ifelse(`Indicator Category` %in% hesitant_group,1,0)) |>
    group_by(`Group Category`,hesitant) |>
    summarise(hesitance_rate = sum(`Estimate (%)`),
              sample_size = first(`Sample Size`)) |>
    mutate(n_hesitance = round(hesitance_rate/100 * sample_size,0)) |>
    filter(hesitant == 1) |>
    select(`Group Category`,n_hesitance,sample_size)
  
  if(nrow(temp) > 0) {
    expected_proportion <- sum(temp$n_hesitance)/sum(temp$sample_size)

  temp <- temp |>
    mutate(expected = sample_size*expected_proportion)
  
  res = chisq.test(x = temp$n_hesitance, p = temp$expected/sum(temp$expected))
  output <- list(round(res$statistic,digits = 0), res$p.value)
    
  }
  else {
    output <- list(0,0)
    
  }
  
  
   return(output)
}

group_name_toloop <- c("Sex","Age","Race/Ethnicity (7 level)","Age by race/ethnicity","Sexual orientation","Gender identity","Metropolitan statistical area","Born in the U.S.", "Language of interview","Poverty status", "Insurance status","Social Vulnerability Index (SVI) of county of residence", "Political leaning of county of residence", "Received non-COVID-19 vaccine(s) in past two years", "Health condition associated with higher risk for COVID-19 (any)","Disability status (any)", "Pregnancy status (females age 18 – 49 years)", "Ever had COVID-19 disease (self-report)","Concern about getting COVID-19 disease", "Confidence in COVID-19 vaccine safety",  "Confidence that COVID-19 vaccine is important", "Healthcare provider recommended I get a COVID-19 vaccine",  "COVID-19 vaccination status of friends and family","Work or school requires COVID-19 vaccination" )

group_name_toloop_short <- lapply(group_name_toloop, function(x) ifelse(x %in% names(group_name_shortening), group_name_shortening[[x]], x))
  

result_df2 <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(result_df2) <- c("group_name", "X2_test_stat", "p_value")

for (x in group_name_toloop_short) {
  res <- get_chi_sq(covid_2021_analysis,geography = "National",
                    time_period = "2021 analysis period",
                    year = 2021,
                    group_name = x,
                    hesitant_group = c("Probably or definitely will not get vaccinated"))
  result_df2[nrow(result_df2) + 1,] = c(x,res[1],res[2])

}

result_df2 <- arrange(result_df2, desc(X2_test_stat))


```

The top 10 most statistically significant group names are shown below, which by the definition of the chi squared test, showed the greatest divergence between the observed and expected values:

```{r}
result_df2 |> ungroup() |> top_n(10,X2_test_stat)

```
Two illustrative mosaic plots are shown below to confirm the chi square test result above - one for the group name with the highest Chi-Square score (the segment that felt vaccines were important), and one for the lowest (gender identity). 


```{r}

mosaic_2var <- function(df,indep_var_group_name,n_chars) {
library(vcd)
  truncate_string <- function(x) {
  str_sub(x, 1, n_chars)}

indep_var_group_name_no_spaces <- gsub("[()\\[/\\\\ ]", ".", indep_var_group_name)
indep_var_group_name_no_spaces <- truncate_string(indep_var_group_name_no_spaces)

temp = df |> ungroup() |> filter(`Group Name` == indep_var_group_name) |> mutate(n_not_hesitant = `Sample Size` - n_hesitant) |>  select(`Group Category`, n_hesitant, n_not_hesitant,`Sample Size`) |>  rename(
    !!indep_var_group_name_no_spaces := `Group Category`,hesitant = `n_hesitant`, `not hesitant` = n_not_hesitant) |> 
  pivot_longer(cols = hesitant:`not hesitant`,names_to = "group", values_to = "Freq" )

temp[[indep_var_group_name_no_spaces]] =  substr(temp[[indep_var_group_name_no_spaces]], 1, n_chars)

formula = as.formula(paste("group", "~", indep_var_group_name_no_spaces))

mosaic(formula, direction = c("v", "h"), temp,
       highlighting_fill = c("grey80", "cornflowerblue"),gp_varnames = gpar(fontsize = 10, fontface = 1), gp_labels = gpar(fontsize = 6))
return(temp)
}

mosaic_2var(covid_2021_analysis,"Vaccine.Imprtnt",100)

mosaic_2var(covid_2021_analysis,"G.Identity",100)


```
It can be clearly seen that those that saw a vaccine as important had significantly lower hesitance than those that did not. However, for the group name with the lowest Chi-Square score (gender identity), the observed plot are exactly as one would expect in a mosaic plot of expected values, thus confirming the statistically insignificant Chi-Square test result.



## Part 2: Analyze which segments showed the greatest reduction in covid vaccine hesitancy between 2021 and 2023


We have already illustrated that national vaccine hesitance declined from around 18% in April/May 2021 to around 11% in May/June 2023. 
```{r}

# Filter for the vaccination uptake and intent indicator
# Filter for responses corresponding to vaccine hesitancy
# Include all adults 18+
# Set time period to monthly (excluding weekly samples) and filter for nationwide samples
vacc_intention_time <- covid |>
  filter(`Indicator Name` == 'Vaccination uptake and intention') |>
  filter(`Indicator Category` == 'Probably or definitely will not get vaccinated') |>
  filter(`Group Name` == 'All adults 18+') |>
  filter(`Time Type` == 'Monthly') |>
  filter(Geography == 'National') |>
  arrange(`Geography Type`, Geography, `Time Period`)

# Derive month of data collection
vacc_intention_time <- vacc_intention_time |>
  mutate(month_split = paste0(str_split(`Time Period`, "-",
                                        simplify = TRUE)[, 2],
         ", ", Year),
         month = mdy(month_split))

# Plot hesitancy over time
ggplot(vacc_intention_time, aes(x = month, y = `Estimate (%)`,
                                color = `Indicator Category`)) +
  geom_line() +
  labs(title = "U.S. Vaccination Response Rates Over Time",
       x = "Month",
       y = "Response Rate (%)") +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Indicator Category:"))


```
However, some segments of the population demonstrated more decline in hesitance than others. The Cleveland dot plot below shows the change in hesitance by population segment, for the 20 most hesitant segments in 2021.

```{r}

annual_monthly_snapshots = c("April 22 - May 29","May 1 - May 28",
                             "April 30 - May 27")

snapshots <- covid |>
  filter(`Time Period` %in% annual_monthly_snapshots,
         `Group Name` %in% group_name_toloop,
         Geography == "National",
         `Indicator Category` == "Probably or definitely will not get vaccinated" ) |>
  mutate(Year = as.factor(Year)) |>
  mutate_at(vars(`Group Name`), ~recode(., !!!group_name_shortening)) |>
  mutate_at(vars(`Group Category`), ~recode(., !!!group_cat_shortening)) |>
  mutate(rows = paste(`Group Name`, `Group Category`, sep = " - ")) |>
  select(rows,Year,`Estimate (%)`) |>
  arrange(rows, Year) 

row_order <- snapshots |>
  pivot_wider(names_from = Year, values_from = `Estimate (%)`) |>
  drop_na(`2021`) |>
  drop_na(`2023`) |>
  mutate(decline_21_23 = `2021` - `2023`) |>
  arrange((decline_21_23))

row_order_list_chg_21_23 <- c(row_order$rows)

row_order_list_2021 <- row_order |>
  arrange(`2021`) |>
  top_n(20,`2021`) |>
  select(rows)

row_order_list_2021 <- row_order_list_2021$rows               

row_order_list_2023 <- c(row_order |>
                           arrange((`2023`)) |>
                           select(rows))

snapshots <- snapshots[snapshots$rows %in% row_order_list_chg_21_23, ]


snapshots |> filter(rows %in% row_order_list_2021) |>
  ggplot(aes(x = `Estimate (%)`, y = fct_relevel(rows,(row_order_list_2021)),
             color = Year)) +
  geom_point(size = 1) +
  labs(title = "20 most hesitant segments in 2021, and changes in 2022-23",
       x = "Vaccine hesitance (%)", y = "Sample segments") +
  theme(legend.position = "bottom",axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 6), plot.title = element_text(size = 10))

```
It can be seen from the Cleveland dot plot that the change in hesitance of the top 20 most hesitant segments in 2021 varies by population segment. To visualise this change better, we created the bar graph below, which shows the 10 segments that demonstrated the largest changes in hesitance between 2021-23.

```{r}
#barplot of change

row_order$rows <- str_wrap(row_order$rows, width = 10)

row_order |>
  top_n (10, decline_21_23) |>
  ggplot(aes(x = fct_reorder(rows,rev(decline_21_23)), y = decline_21_23)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 10 categories with the largest 2021-23 changes",
       x = "Group categories", y = "2021-23 change in hesitance") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 0, vjust = 1, hjust=0.5,size=6),
        axis.text.y = element_text(size=8),
        axis.title = element_text(size = 8),
        plot.title = element_text(size = 8))
```

The bar chart shows the largest increases in  segments that believed vaccines were a little or not at all important, followed by those that had low confidence in vaccine safety/efficacy, and those trying to get pregnant.

The scattergraph below shows the relationship between initial hesitance in 2021 and change in hesitance between 2021-23 across population segments: segments with the highest initial hesitance generally showed the greatest reduction in hesitance, with a high R2 of 0.75. This could be because those with the highest hesitance received targeted public healthcare interventions, or had the greatest scope for reduction in the first place - determining the exact reason requires further investigation that is beyond the scope of this project.


```{r}

row_order %>% ggplot(aes(x = `2021`, y = decline_21_23)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Scatterplot of 2021 hesitance vs. 2021-23 decline by population segment",
       x = "2021 Hesitance (%)",
       y = "Decline in hesitance (2021-23) (%)")+
  stat_poly_eq(formula = y ~ x,   # Add R-squared value
               aes(label = paste( ..rr.label.., sep = "~~~~")),
               parse = TRUE,
               size = 4)
```


To understand changes in hesitance by geography, we visualised the change in hesitance by state in the spatial heatmap below:
```{r}

comparison_baseline_2021 <- covid |>
  filter(`Year` == 2021, `Time Type` == "Monthly",
         `Time Period` == "April 22 - May 29",
         `Group Category`== "All adults age 18+ years",
         `Geography Type` %in% c("Jurisdictional Estimates"),
         `Indicator Category` == "Probably or definitely will not get vaccinated",
         !grepl("-", Geography)) |>
  mutate(n_hesitant = round(`Estimate (%)`/100 * `Sample Size`, 0)) |>
  group_by(Geography) |>
  summarise(`Sample Size` = sum(`Sample Size`),
            n_hesitant = sum(n_hesitant)) |>
  mutate(`Estimate (%)` = n_hesitant / `Sample Size` * 100,
         `Time Period` = "2021 most recent snapshot",
         Year = 2021)

covid_2023_analysis_allstates <- covid |>
  filter(`Year` == 2023, `Time Type` == "Monthly",
         `Time Period` == "May 28 - June 30",
         `Group Category`== "All adults age 18+ years",
         `Geography Type` %in% c("Jurisdictional Estimates"),
         `Indicator Category` == "Probably or definitely will not get vaccinated",
         !grepl("-", Geography)) |>
  mutate(n_hesitant = round(`Estimate (%)`/100 * `Sample Size`, 0)) |>
  group_by(Geography) |>
  summarise(`Sample Size` = sum(`Sample Size`),
            n_hesitant = sum(n_hesitant)) |>
  mutate(`Estimate (%)` = n_hesitant / `Sample Size` * 100,
         `Time Period` = "2023 most recent snapshot",
         Year = 2023)


state_delta <- left_join(covid_2023_analysis_allstates,
                         comparison_baseline_2021 |>
                           select(Geography, estimate_2021 = `Estimate (%)`),
                         by = "Geography")

state_delta <- state_delta |>
  mutate(`Change in Hesitancy Rates` = `Estimate (%)` - estimate_2021)

state_delta_map <- state_delta |>
  select(Geography, `Change in Hesitancy Rates`) |>
  transmute(region = tolower(Geography),
            value = as.numeric(`Change in Hesitancy Rates`))


# red_green_palette <- brewer.pal(n = 7, name = "RdYlGn")

state_choropleth(state_delta_map,
                 title = "   2021-2023 Changes in Vaccine Hesitance by State",
                 legend = "Change in \nHesitancy Rates   ",
                 num_colors = 0)


```

The spatial heatmap indicates that West Virginia exhibited the greatest overall change in hesitance (23.3%), while New Mexico experienced a slight increase in hesitance of 0.3%. 

The scattergraph below shows the relationship between initial hesitance in 2021 and change in hesitance between 2021-23 by state: states with the highest initial hesitance generally showed the greatest decline in hesitance, although the R2 is low (0.15). The relationship is significantly weaker than a similar relationship seen for population segments, where the R2 was 0.75.


```{r}

state_delta %>% mutate(`Decline in Hesitancy Rates` = -1*`Change in Hesitancy Rates`) %>% ggplot(aes(x = estimate_2021, y = `Decline in Hesitancy Rates`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(title = "Scatterplot of 2021 hesitance vs. 2021-23 decline by state",
       x = "2021 Hesitance (%)",
       y = "Decline in hesitance (2021-23) (%)")+
  stat_poly_eq(formula = y ~ x,   # Add R-squared value
               aes(label = paste( ..rr.label.., sep = "~~~~")),
               parse = TRUE,
               size = 4)


```




```{r}

hesitance_state_group_cat <- covid_hes_prepped |>
  filter(`Geography Type` %in% c("Jurisdictional Estimates"),
         !grepl("-", Geography)) |>
  mutate(month_split = paste0(str_split(`Time Period`, "-",
                                        simplify = TRUE)[, 2], ", ", Year),
         month = mdy(month_split),
         n_hesitant = round(`Estimate (%)`/100 * `Sample Size`, 0)) |>
  group_by(Geography, `Group Name`, `Group Category`, concat, month) |>
  summarise(`Sample Size` = sum(`Sample Size`),
            n_hesitant = sum(n_hesitant)) |>
  mutate(`Estimate (%)` = n_hesitant / `Sample Size` * 100,
         concat_no_state = concat,
         concat = paste0(Geography, ": ", concat)) |>
  ungroup()


d3_input_prep <- hesitance_state_group_cat |>
  pivot_wider(id_cols = month, names_from = "concat", values_from = `Estimate (%)`) |>
  dplyr::rename(Category = month)

d3_input_prep[is.na(d3_input_prep)] <- 0


write_csv(d3_input_prep, "D3 Inputs/All State-Category Values.csv")


dropdown_selections <- hesitance_state_group_cat |>
  distinct(Geography, concat_no_state)

write_csv(dropdown_selections, "D3 Inputs/Dropdown Choices.csv")

```



## Part 3: Based on the findings above, we will propose high-level recommendations for segments of the population that need further public health interventions.

Our project highlighted 3 clear trends:

1. Statistically significant drivers to covid vaccine hesitance in 2021 were those that had reservations about the importance or safety and efficacy of vaccines, and those with unvaccinated friends and family.
2. Hesitance can be reduced - there was a history of successful reductions in hesitance between 2021-23 nationally (18%), state-wise (up to 23%) and population segment-wise (32%). This was especially visible in states like W Virginia, and segments such as those who had reservations about the importance or the safety and efficacy of vaccines. This indicates potential to further reduce hesitance in these segments, although further study into the causes of these declines in hesitance is required.
3. 3 years after the beginning of the pandemic, vaccine hesitance persists in some segments of the populations, and in some states, and within vulnerable population segments such as pregnant women and those with unvaccinated family members

Our findings suggest the need for further study into the drivers of reduced hesitance in the population segments mentioned above, to explore what additional public healthcare interventions would further reduce hesitance, especially among vulnerable demographics.
